---
title: "TE_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# **Example set-up: The user can revise this section**
## **This section illustrates use of the functions**
Additional documentation of the main functions at the top of the **TEMAP2_functions.R**


```{r}

library(readxl)
library(boot)
library(scales)
library(data.table)
library(ggplot2)
```



##    **Read in the data**
```{r}

setwd("C:/Users/edika/Desktop/key backup/Michael") #working directory
source("TEMAP2_only_functions.R")

data_file<-"Example.xlsx"            # Input data are in a xlsx file

read_excel(data_file, sheet = "Inputs")[, c("Inputs", "Specifications")]
read_excel(data_file, sheet = "Participant responses")

response_patterns <-read_excel(data_file, sheet = "Participant responses")
number_cases <- nrow(response_patterns)

```

```{r error=TRUE}

simulation_results_list <-TE_READ_DATA(data_file)
names(simulation_results_list)
(simulation_results_list[[3]])

output_file<- paste(simulation_results_list$Output_file,".txt", sep = "")
(fit_method <- simulation_results_list$fit_method)

#   It may take 20 min. per case to simulate TE model and analyze 10,000 samples via MC


#   Output is directed to the file designated below

#sink(output_file,append=TRUE)
label_TE_4_errors=c("OUTPUT >>> True and Error Model Results - 4 error terms >>>")
label_TE_2_errors=c("OUTPUT >>> True and Error Model Results - 2 error terms >>>")
label_TE_1_errors=c("OUTPUT >>> True and Error Model Results - 1 error term >>>")
label_Indep=c("OUTPUT >>> Independence Model Results")

print(label_TE_4_errors)
print(simulation_results_list$results_TE_4_errors)
print(label_TE_2_errors)
print(simulation_results_list$results_TE_2_errors)
print(label_TE_1_errors)
print(simulation_results_list$results_TE_1_errors)
print(label_Indep)
print(simulation_results_list$results_Independence)

print("OUTPUT >>> First few samples parameters for Case 1, Model TE-4 >>>") 
print(head(simulation_results_list$results_TE_4_MC[[1]]$parameters_Re.fitted.MC))
print(head(simulation_results_list$results_TE_4_MC[[1]]$errors_Re.fitted.MC))

print("OUTPUT >>> First few Monte Carlo Re.fitted Chi-squared (or G2) values for Case 1, Model TE-4 >>>") 
print(head(simulation_results_list$results_TE_4_MC[[1]]$chisq_Re.fitted.MC))

print("OUTPUT >>> Chi-squared (or G2) difference test for Case 1, for selected models (available when the Monte Carlo test was not selected) >>>") 
print(simulation_results_list$results_Chi2_diff_table[[1]]$TE_1_Chi2_diff_test)
print(simulation_results_list$results_Chi2_diff_table[[1]]$TE_2_Chi2_diff_test)
print(simulation_results_list$results_Chi2_diff_table[[1]]$TE_4_Chi2_diff_test)

print("OUTPUT >>> Monte Carlo Chi-squared (or G2) difference test for Case 1, for selected models >>>") 
print(simulation_results_list$results_Chi2_diff_MC[[1]]$TE_1_Chi2_diff_test)
print(simulation_results_list$results_Chi2_diff_MC[[1]]$TE_2_Chi2_diff_test)
print(simulation_results_list$results_Chi2_diff_MC[[1]]$TE_4_Chi2_diff_test)

#sink(file = NULL)

```

##  **Draw histogram of the fit for the TE-4 Model**

##   Draw histogram for **Case #1** TE-4 Model **Refit** to each sample
Change to 'results_TE_2_MC' for TE-2 model or change to 'results_TE_1_MC' for TE-1 model.
```{r}

case=1
if (is.null(simulation_results_list$results_TE_4_MC)==0) {
  #pdf("MonteCarlo_TE_4_Refit_Case_1.pdf")
  MonteCarlo_statistic_histogram <- hist(simulation_results_list$results_TE_4_MC[[case]]$chisq_Re.fitted.MC,col=scales::alpha('skyblue',.5),
                                         breaks=50,xlim = c(0, simulation_results_list$results_TE_4_MC[[case]]$original.chisq[[1]]+10),border=F,
                                         main=paste0("Monte Carlo TE - 4 Model (Re-Fit) - Case ", case),
                                         xlab=paste0(fit_method," Re-Fit in each Sample")) 
  abline(v=simulation_results_list$results_TE_4_MC[[case]]$original.chisq,col="red")
  #dev.off()

}
```

## Draw histogram for **Case #1** TE Model **Conservative** Monte Carlo
  
```{r}

if (is.null(simulation_results_list$results_TE_4_MC)==0) {

  
  # pdf("MonteCarlo_TE_4_Conservative_Case_1.pdf")
  MonteCarlo_conservative_statistic_histogram <- hist(simulation_results_list$results_TE_4_MC[[case]]$chisq_Conservative.MC,col=scales::alpha('skyblue',.5),
                                                      breaks=50,
                                                      xlim = c(0, simulation_results_list$results_TE_4_MC[[case]]$original.chisq[[1]] +10),
                                                      border=F,
                                                      main=paste0("Monte Carlo TE - 4 Model (Conservative) - Case ", case),
                                                      xlab=paste0(fit_method," Conservative"))
  abline(v=simulation_results_list$results_TE_4_MC[[case]]$original.chisq,col="red")
  #dev.off()
}

```

##  **Draw histogram of the fit for the TE-2 Model**

##  Draw Histogram for **Case #1** TE-2 Model **Refit** to each sample
Change to 'results_TE_2_MC' for TE-2 model or change to 'results_TE_1_MC' for TE-1 model.
```{r}

if (is.null(simulation_results_list$results_TE_2_MC)==0) {
  #pdf("MonteCarlo_TE_2_Refit_Case_1.pdf")
  MonteCarlo_statistic_histogram <- hist(simulation_results_list$results_TE_2_MC[[case]]$chisq_Re.fitted.MC,col=scales::alpha('skyblue',.5),
                                         breaks=50,
                                         xlim = c(0, simulation_results_list$results_TE_2_MC[[case]]$original.chisq[[1]]+10),
                                         border=F,
                                         main=paste0("Monte Carlo TE - 2 Model (Re-Fit) - Case ", case),
                                         xlab=paste0(fit_method," Re-Fit in each Sample")) 
  abline(v=simulation_results_list$results_TE_2_MC[[case]]$original.chisq,col="red")
  #dev.off()
}
```
## Draw Histogram for **Case #1** TE Model **Conservative** Monte Carlo
  
```{r}
  if (is.null(simulation_results_list$results_TE_2_MC)==0) {

  #pdf("MonteCarlo_TE_2_Conservative_Case_1.pdf")
  MonteCarlo_conservative_statistic_histogram <- hist(simulation_results_list$results_TE_2_MC[[case]]$chisq_Conservative.MC,col=scales::alpha('skyblue',.5),
                                                      breaks=50,
                                                      xlim = c(0, simulation_results_list$results_TE_2_MC[[case]]$original.chisq[[1]]+10),
                                                      border=F,
                                                      main=paste0("Monte Carlo TE - 2 Model (Conservative) - Case ", case),
                                                      xlab=paste0(fit_method," Conservative")) 
  abline(v=simulation_results_list$results_TE_2_MC[[case]]$original.chisq,col="red")
  #dev.off()
}

```


##  **Save bootstrapped density GRAPHS for the TE-4 Model**

The next section shows how to graph the results of the bootstrapping distributions for Case #1.

Here smoothed density distributions are drawn.  One could use histograms instead.

Note that parameters fixed by the user do not show density distributions.

Change to 'rresults_TE_2_Boot_CI' for TE-2 model or change to 'results_TE_1_Boot_CI' for TE-1 model.

```{r}

if (is.null(simulation_results_list$results_TE_4_Boot_CI)==0) {
  #pdf("Bootstrapped Density of a_00.pdf")
  plot(density( simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.parameters[,c("a_00")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_00, TE - 4, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("Bootstrapped Density of a_01.pdf")
  plot(density(simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.parameters[,c("a_01")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_01, TE - 4, Fit method: ", fit_method, ", Case: ",case))                                               
  #dev.off()
  #pdf("Bootstrapped Density of a_10.pdf")
  plot(density(simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.parameters[,c("a_10")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_10, TE - 4, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("Bootstrapped Density of a_11.pdf")
  plot(density(simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.parameters[,c("a_11")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_11, TE - 4, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("Bootstrapped Density of error_R1.pdf")
  plot(density(simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.errors[,c("error_R1")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of error_R1, TE - 4, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("Bootstrapped Density of error_R2.pdf")
  plot(density(simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.errors[,c("error_R2")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of error_R2, TE - 4, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("Bootstrapped Density of error_S1.pdf")
  plot(density(simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.errors[,c("error_S1")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of error_S1, TE - 4, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("Bootstrapped Density of error_S2.pdf")
  plot(density(simulation_results_list$results_TE_4_Boot_CI[[case]]$boot.errors[,c("error_S2")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of error_S2, TE - 4, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
}

```


## **Save bootstrapped density graphs for the TE-2 Model**

The next section shows how to graph the results of the bootstrapping distributions for Case #1

Here smoothed density distributions are drawn.  One could use histograms instead.

Note that parameters fixed by the user do not show density distributions 

Change to 'rresults_TE_2_Boot_CI' for TE-2 model or change to 'results_TE_1_Boot_CI' for TE-1 model.

```{r}
#dev.set(dev.next())
#dev.cur()

if (is.null(simulation_results_list$results_TE_2_Boot_CI)==0) {
  
  #pdf("TE-2 Bootstrapped Density of a_00.pdf")
  plot(( simulation_results_list$results_TE_2_Boot_CI[[1]]$boot.parameters[,c("a_00")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_00, TE - 2, Fit method: ", fit_method, ", Case: ",case))
  # dev.off()
  #pdf("TE-2 Bootstrapped Density of a_01.pdf")
  plot(density(simulation_results_list$results_TE_2_Boot_CI[[1]]$boot.parameters[,c("a_01")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_01, TE - 2, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("TE-2 Bootstrapped Density of a_10.pdf")
  plot(density(simulation_results_list$results_TE_2_Boot_CI[[1]]$boot.parameters[,c("a_10")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_10, TE - 2, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("TE-2 Bootstrapped Density of a_11.pdf")
  plot(density(simulation_results_list$results_TE_2_Boot_CI[[1]]$boot.parameters[,c("a_11")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of a_11, TE - 2, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("TE-2 Bootstrapped Density of error_1.pdf")
  plot(density(simulation_results_list$results_TE_2_Boot_CI[[1]]$boot.errors[,c("error_1")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of error_1, TE - 2, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  #pdf("TE-2 Bootstrapped Density of error_2.pdf")
  plot(density(simulation_results_list$results_TE_2_Boot_CI[[1]]$boot.errors[,c("error_2")]),xlim = c(0, 1), main=paste0("Bootstrapped Density of error_2, TE - 2, Fit method: ", fit_method, ", Case: ",case))
  #dev.off()
  
}

```

## **Draw histograms of the Chi-squared difference test for the restricted and unrestricted models**

## Draw Histogram for **Case #1** of the Chi-squared difference values for the TE-4 Model
Change participant in '[[case]]'. 
```{r}

TE_4_Re_fit <- simulation_results_list$results_Chi2_diff_MC[[case]]$TE_4_Chi2_diff_test$Chi2_diff_Re.fitted.MC$MC_Re.fitted_chi2_diff_results[,3]
TE_2_Re_fit <- simulation_results_list$results_Chi2_diff_MC[[case]]$TE_2_Chi2_diff_test$Chi2_diff_Re.fitted.MC$MC_Re.fitted_chi2_diff_results[,3]
TE_1_Re_fit <- simulation_results_list$results_Chi2_diff_MC[[case]]$TE_1_Chi2_diff_test$Chi2_diff_Re.fitted.MC$MC_Re.fitted_chi2_diff_results[,3]

TE_4_Standard <- simulation_results_list$results_Chi2_diff_MC[[case]]$TE_4_Chi2_diff_test$Chi2_diff_standard[,3]
TE_2_Standard <- simulation_results_list$results_Chi2_diff_MC[[case]]$TE_2_Chi2_diff_test$Chi2_diff_standard[,3]
TE_1_Standard <- simulation_results_list$results_Chi2_diff_MC[[case]]$TE_1_Chi2_diff_test$Chi2_diff_standard[,3]
```


Adjust the number of breaks for a better display

Negative differences in Chi2 could occur in rare occasions, only when the optimization algorithm reaches local minima points.

```{r}

if (is.null(TE_4_Re_fit)==0) {
  #pdf(paste0("MonteCarlo_",fit_method,"_difference_test_TE_4_Refit_Case_",case, ".pdf"))
  MonteCarlo_histogram <- hist(TE_4_Re_fit,col=scales::alpha('skyblue',.5),breaks=30,
                               border=F,
                               main=paste0("Monte Carlo ",fit_method," Difference, TE - 4 Model (Re-Fit)"),
                               xlim = c(-1, TE_4_Standard+4),
                               xlab=paste0(fit_method," Difference (Re-Fit Monte Carlo)"))
  abline(v=TE_4_Standard,col="red")
  #dev.off()
}  

if (is.null(TE_2_Re_fit)==0) {
  #pdf(paste0("MonteCarlo_",fit_method,"_difference_test_TE_2_Refit_Case_",case, ".pdf"))
  MonteCarlo_histogram <- hist(TE_2_Re_fit,col=scales::alpha('skyblue',.5),breaks=30,
                               border=F,
                               main=paste0("Monte Carlo ",fit_method," Difference, TE - 2 Model (Re-Fit)"),
                               xlim = c(-1, TE_2_Standard+4),
                               xlab=paste0(fit_method," Difference (Re-Fit Monte Carlo)"))
  abline(v=TE_2_Standard,col="red")
  #dev.off()
}  

```


This example is set up to save only selected results to illustrate the program

For example, it saves only graphs for the first case.

You can easily revise it to save other information and to make other graphs  

See documentation section.

# **For profesoor Lee**

```{r}

fit_method <- simulation_results_list$fit_method
if (is.null(simulation_results_list$results_TE_4_Boot_CI)==0) {
  
  file_name <- paste0("Boot sample estimates, TE4, fit method ", fit_method,
                      " restrictions a_00=",simulation_results_list$user_restrictions[1],  
                      " ,a_01=",simulation_results_list$user_restrictions[2],
                      " ,a_10=",simulation_results_list$user_restrictions[3],
                      " ,a_11=",simulation_results_list$user_restrictions[4],
                      ".csv")
  
  boot_estimates_all <- NULL
  for (i in 1:length(simulation_results_list$results_TE_4_Boot_CI)) {
    boot_estimates<- as.data.table(simulation_results_list$results_TE_4_Boot_CI[[i]]$boot.parameters)
    boot_estimates <- cbind(boot_estimates,simulation_results_list$results_TE_4_Boot_CI[[i]]$boot.errors)
    
    boot_estimates[,case:=i]
    
    boot_estimates_all <- rbind(boot_estimates_all, boot_estimates)
  }
  
  write.csv(boot_estimates_all, file= file_name)
  
  
}


if (is.null(simulation_results_list$results_TE_2_Boot_CI)==0) {
  
  
  file_name <- paste0("Boot sample estimates, TE2, fit method ", fit_method, " restrictions a_00=",simulation_results_list$user_restrictions[1],  
                      " ,a_01=",simulation_results_list$user_restrictions[2],
                      " ,a_10=",simulation_results_list$user_restrictions[3],
                      " ,a_11=",simulation_results_list$user_restrictions[4],
                      ".csv")
  
  boot_estimates_all <- NULL
  for (i in 1:length(simulation_results_list$results_TE_2_Boot_CI)) {
    boot_estimates<- as.data.table(simulation_results_list$results_TE_2_Boot_CI[[i]]$boot.parameters)
    boot_estimates <- cbind(boot_estimates,simulation_results_list$results_TE_2_Boot_CI[[i]]$boot.errors)
    
    boot_estimates[,case:=i]
    
    boot_estimates_all <- rbind(boot_estimates_all, boot_estimates)
  }
  
  write.csv(boot_estimates_all, file= file_name)
  
}



if (is.null(simulation_results_list$results_TE_1_Boot_CI)==0) {
  
  
  file_name <- paste0("Boot sample estimates, TE1, fit method ", fit_method, " restrictions a_00=",simulation_results_list$user_restrictions[1],  
                      " ,a_01=",simulation_results_list$user_restrictions[2],
                      " ,a_10=",simulation_results_list$user_restrictions[3],
                      " ,a_11=",simulation_results_list$user_restrictions[4],
                      ".csv")
  
  boot_estimates_all <- NULL
  for (i in 1:length(simulation_results_list$results_TE_1_Boot_CI)) {
    boot_estimates<- as.data.table(simulation_results_list$results_TE_1_Boot_CI[[i]]$boot.parameters)
    boot_estimates <- cbind(boot_estimates,simulation_results_list$results_TE_1_Boot_CI[[i]]$boot.errors)
    
    boot_estimates[,case:=i]
    
    boot_estimates_all <- rbind(boot_estimates_all, boot_estimates)
  }
  
  write.csv(boot_estimates_all, file= file_name)
  
}

```

# Plot of confidence intervals

Confidence intervals are ommited for the parameters constrained to be 0 in the input file

```{r}

case=1 #select case for graph

CI_plot_TE1 <-NULL
CI_plot_TE2 <-NULL
CI_plot_TE4 <-NULL

if (is.null(simulation_results_list$results_TE_1_Boot_CI)==0) {
  
  CI_plot_TE1<-(simulation_results_list$results_TE_1_Boot_CI[[case]]$confidence.intervals)
  CI_plot_TE1 <-cbind(Parameters=rownames(CI_plot_TE1), CI_plot_TE1)
  CI_plot_TE1 <- as.data.table(CI_plot_TE1)
  CI_plot_TE1[,Estimates:=as.numeric(Estimates)]
  CI_plot_TE1[,`Lower bound`:=as.numeric(`Lower bound`)]
  CI_plot_TE1[,`Upper bound`:=as.numeric(`Upper bound`)]
  CI_plot_TE1[,Model:= "TE1"]
}

if (is.null(simulation_results_list$results_TE_2_Boot_CI)==0) {
  
  
  CI_plot_TE2<-(simulation_results_list$results_TE_2_Boot_CI[[case]]$confidence.intervals)
  CI_plot_TE2<-cbind(Parameters=rownames(CI_plot_TE2), CI_plot_TE2)
  CI_plot_TE2 <- as.data.table(CI_plot_TE2)
  CI_plot_TE2[,Estimates:=as.numeric(Estimates)]
  CI_plot_TE2[,`Lower bound`:=as.numeric(`Lower bound`)]
  CI_plot_TE2[,`Upper bound`:=as.numeric(`Upper bound`)]
  CI_plot_TE2[,Model:= "TE2"]
}

if (is.null(simulation_results_list$results_TE_4_Boot_CI)==0) {
  
  
  CI_plot_TE4<-(simulation_results_list$results_TE_4_Boot_CI[[case]]$confidence.intervals)
  CI_plot_TE4<-cbind(Parameters=rownames(CI_plot_TE4), CI_plot_TE4)
  CI_plot_TE4 <- as.data.table(CI_plot_TE4)
  CI_plot_TE4[,Estimates:=as.numeric(Estimates)]
  CI_plot_TE4[,`Lower bound`:=as.numeric(`Lower bound`)]
  CI_plot_TE4[,`Upper bound`:=as.numeric(`Upper bound`)]
  CI_plot_TE4[,Model:= "TE4"]
}

CI_plot_all <- rbind(CI_plot_TE1, CI_plot_TE2, CI_plot_TE4)

# Make the graph with the 95% confidence interval
if (is.null(CI_plot_all)!=1) {
  
  figure <- ggplot(CI_plot_all, aes(x=Parameters, 
                                    y=Estimates #, group=1
                                    #ymin=CI_plot$`Lower bound`, ymax=CI_plot$`Upper bound`
  )) +
    geom_errorbar(width=.1, aes(ymin=CI_plot_all$`Lower bound`, ymax=CI_plot_all$`Upper bound`)) +
    geom_point(shape=21, size=3, fill="white")  +  coord_flip() +  
    facet_grid( Model~., scales = "free") +
    ggtitle(paste0("Bootstrap CI for TE models; Case ",case," ; Fit method: ", fit_method,
                   ";\nRestrictions: a_00=",simulation_results_list$user_restrictions[1],  
                   ", a_01=",simulation_results_list$user_restrictions[2],
                   ", a_10=",simulation_results_list$user_restrictions[3],
                   ", a_11=",simulation_results_list$user_restrictions[4]
    ))
  ggsave(figure, file= paste0("Bootstrap CI for TE models; Case ",case," ; Fit method= ", fit_method,
                              "; Restrictions a_00=",simulation_results_list$user_restrictions[1],  
                              ", a_01=",simulation_results_list$user_restrictions[2],
                              ", a_10=",simulation_results_list$user_restrictions[3],
                              ", a_11=",simulation_results_list$user_restrictions[4], ".pdf"  ), width=7,height=8)
  figure
}


```


